Open Preprints in Ecology \& Evolution

### 2012-09-03 ms.tex
main text

@@

%% Essentially an open equivalent to sharing preprints between colleagues.

@@

@@

While the practice is still rare, preprints are becoming more common in
biological sciences, which is experiencing faster growth in arXiv submission
than any other fields \cite{cal12}. Also, most scientific journals are
preprint-friendly: Nature, PLOS, BMC, PNAS, Science (mostly)
\ref{table:policies}, and all the journals from Elsevier and Springer. Very
recently, the Ecological Society of America recently changed its policy to allow
public preprints (REF). In our field, very few scientific publications will not
consider a manuscript submitted to arXiv. Many ecology \& evolution journals
adopt a ``by default'' hostile attitude towards preprints, mostly due to the
lack of clear policy of the publishers. As an example, Wiley-Blackwell, which
publishes some of the leading journal in the field, has no official policy on
the subject.

@@

% Short paragraphs on the current open preprint repositories.

@@

arXiv (\href{http://arxiv.org/}{http://arxiv.org/})

arXiv is funded by a network of universities.

@@

@@

% Ethan P. White's section

@@

% Not really the same, but worth mentionning.

>>>>

Preprints servers such as arXiv are common in mathematics and physics, but still
a minority of papers in ecology and evolution use them. Preprints servers allow
author to make their manuscripts publicly available before submitting them to
journals for peer-review. ... The idea became popular with arXiv, and many
physicists start their day with a look at the new papers in arXiv \cite{gin11}

@@

@@

Some of the responses to public preprints are surprisingly since they are,
essentially, the same as exchanging preprints among colleagues. Prepublication
reviews by a small network of colleagues is an important part of the scientific
process. Preprints servers simply offer a way to extend this network of
colleagues to the entire scientific community. It ensures that science is not
constrained by small networks of scientists exchanging ideas. Ginsparg made
arXiv.org in part for democratic reasons: he wanted everyone from graduate
students in small universities to Princeton professors to have access to the
most recent scientific ideas. Ginsparg revolutionary idea was simply to use the
power of the internet for preprints, not just for the end product, so the
process can be open from A to Z, instead of being just open at the end of the
process.

@@

While the practice is still rare, preprints are becoming more common in
biological sciences, which is experiencing faster growth in arXiv submissions
than any other fields \cite{cal12}. Also, most scientific journals are
preprint-friendly: Nature, PLOS, BMC, PNAS, Science (mostly)
\ref{table:policies}, and all the journals from Elsevier and Springer. Very
recently, the Ecological Society of America recently changed its policy to allow
public preprints (REF). In our field, few scientific publications will not
consider a manuscript submitted to arXiv. Still, many ecology \& evolution journals
adopt a ``by default'' hostile attitude towards preprints, mostly due to the
lack of clear policy of the publishers. As an example, Wiley-Blackwell, which
publishes some of the leading journal in the field, has no official policy on
the subject \ref{table:policies}.

@@

We briefly discuss the main options to submit preprints to open servers:
arXiv.org, Figshare, and the upcoming PeerJ and F1000Research.

@@

arXiv (\href{http://arxiv.org/}{http://arxiv.org/}).
arXiv is funded by a network of universities.

% @Joel

@@

% @PhDP

@@

% @Ethan

@@

% Written from memory: to verify!

F1000Research is not a public preprint server like the previous three servers.
Whereas arXiv, Figshare, and PeerJ offer an option to submit a manuscript
without having it reviewed, papers submitted to F1000Research will eventually be
reviewed. Thus, F1000Research offers a hybrid model with publicly available
manuscripts at time of submission and standard peer-reviews. Manuscripts are
considered ``accepted'' and will only be indexed after two positive referee
response.
